\textbf{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}

### 2020-10-22 arxiv_full_version/diff_cov.tex
polish typo

The previous section assumes that all tasks have the same sample size and feature vectors.
In this section, we study how having different sample sizes and different covariance matrices impact hard parameter sharing.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
The setting where covariates differ across tasks is often called ``covariate shift''.

Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p} \text{ and }
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
We consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \citet{WZR20}).
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where }
	\hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
Our main result in this section characterizes the asymptotic bias-variance limits in the high-dimensional setting.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote $M$'s singular values in descending order.
Our main result is stated as follows.

@@

In the above inequality, the $d^2$ scaling term is the bias limit and the $\sigma^2$ scaling term is the variance limit.
This result allows for a more concrete interpretation since the dependence on datasets' properties is explicit.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.
As a remark, %in equation \eqref{cor_MTL_error}, the predication loss $L(\hat{\beta}_2^{\MTL}) $ was obtained using the global minimizer $(\hat A,\hat B)$. B
by combining the bias and variance limits, we can also obtain a bias-variance tradeoff for any local minimizer of $f(A, B)$.
The proof is similar to Corollary \ref{cor_MTL_loss}, so we omit the details.

Next, we use the bias-variance limits to study how varying sample sizes impacts HPS.
For example, imagine if we want to decide whether to collect more of task one's data or not, how does increasing $n_1$ affect the prediction loss?
We assume that $n_2$ is fixed for simplicity.
%Now we illustrate an interesting phenomenon that adding task one's samples helps task two initially, but may hurt eventually.
The variance limit in equation \eqref{cor_MTL_error} obvious decreases with $n_1$.
It turns out that the bias term always increases with $n_1$, which can be verified by showing that the bias limit's derivative is always nonnegative.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
By comparing the derivative of the bias and variance limits with respect to $n_1$ (details omitted), we obtain the following dichotomy.
\begin{enumerate}
	\item When $\frac{d^2}{\sigma^2} < \frac{p}{4n_2 - 6p}$, the prediction loss decreases monotonically as $n_1$ increases.
	Intuitively, this regime of $d^2$ always helps task two.
	\item When $\frac{d^2}{\sigma^2} > \frac{p}{4n_2 - 6p}$, the prediction loss always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ (when $n_1 = 0$), and then increases to $d^2$ (when $n_1 \rightarrow \infty$).
	To see this, near the point where $n_1$ is zero, one can verify (from the derivatives) that bias increases less while variance decreases more, and there is \textit{exactly} one critical point where the derivative is zero, which corresponds to the \textit{optimal sample size ratio}.
\end{enumerate}

>>>>

The previous section assumes that all tasks have the same sample size and feature vectors.
This section discusses how having different sample sizes and different covariance matrices impact hard parameter sharing.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
The setting where covariates differ across tasks is often known as ``covariate shift''.

Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p} \text{ and }
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Without loss of generality, we consider task two's prediction loss, and the same result applies to task one.
We consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \citet{WZR20}).
When $r = 1$, $B$ is a vector, and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where }
	\hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
Our main result in this section characterizes the asymptotic bias-variance limits in the high-dimensional setting.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a critical quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote $M$'s singular values in descending order.
Our main result is stated as follows.

@@

In the above inequality, the $d^2$ scaling term is the bias limit, and the $\sigma^2$ scaling term is the variance limit.
This result allows for a more concrete interpretation since the dependence on datasets' properties is explicit.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.
As a remark, %in equation \eqref{cor_MTL_error}, the predication loss $L(\hat{\beta}_2^{\MTL}) $ was obtained using the global minimizer $(\hat A,\hat B)$. B
by combining the bias and variance limits, we can also obtain a bias-variance tradeoff for any local minimizer of $f(A, B)$.
The proof is similar to Corollary \ref{cor_MTL_loss}, so we omit the details.

Next, we use the bias-variance limits to study how varying sample sizes impacts HPS.
For example, imagine if we want to decide whether to collect more of task one's data or not, how does increasing $n_1$ affect the prediction loss?
We assume that $n_2$ is fixed for simplicity.
%Now we illustrate an interesting phenomenon that adding task one's samples helps task two initially, but may hurt eventually.
The variance limit in equation \eqref{cor_MTL_error} obviously decreases with $n_1$.
It turns out that the bias term always increases with $n_1$, which can be verified by showing that the bias limit's derivative is always nonnegative.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
By comparing the derivative of the bias and variance limits with respect to $n_1$ (details omitted), we obtain the following dichotomy.
\begin{enumerate}
	\item When $\frac{d^2}{\sigma^2} < \frac{p}{4n_2 - 6p}$, the prediction loss decreases monotonically as $n_1$ increases.
	Intuitively, this regime of $d^2$ always helps task two.
	\item When $\frac{d^2}{\sigma^2} > \frac{p}{4n_2 - 6p}$, the prediction loss always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ (when $n_1 = 0$), and then increases to $d^2$ (when $n_1 \rightarrow \infty$).
	To see this, near the point where $n_1$ is zero, one can verify (from the derivatives) that bias increases less while variance decreases more, and there is \textit{exactly} one critical point where the derivative is zero, which corresponds to the \textit{optimal sample size ratio}.
\end{enumerate}

### 2020-10-22 arxiv_full_version/same_cov.tex
polish typo

%We begin by considering the case where all tasks have the same sample size and feature covariates, that is, $n_i = n$ and $X^{(i)} = X\in\real^{n\times p}$ for all $i = 1, \dots, t$.
%We provide a sharp generalization error bound of hard parameter sharing estimators.
In this section, we show that the prediction loss of hard parameter sharing admits a clean bias-variance decomposition, when all tasks have the same features.
%\FY{Is it should be "same covariate" or "same covariates"?}
%This setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
%We consider an arbitrary local minimum $B, W_1, \dots, W_2$ of the optimization objective.
%We extend the bias-variance decomposition from the two-task case to the multiple-task case.
%We observe that the expected prediction loss of $\hat{\beta}_t^{\MTL}$ conditional on $X$ consists of a bias and a variance equation as follows
%\begin{align}
%	\exarg{\varepsilon_1, \dots, \varepsilon_t}{L(\hat{\beta}_t^{\MTL}) \mid X}
%	=& \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} \cW^{\top} (\cW \cW^{\top})^{-1} W_t - \beta_t}}^2 \label{eq_bias_multiple} \\
%	&+ \sigma^2 \cdot (W_t^{\top} (\cW \cW^{\top})^{-1} W_t) \cdot \bigtr{\Sigma (X^{\top} X)^{-1}} \label{eq_var_multiple}
%\end{align}
%One can see that equation \eqref{eq_bias_multiple} is the bias of the multi-task learning estimator and equation \eqref{eq_var_multiple} is its variance.
%Compared to the prediction loss of single-task learning (cf. equation \eqref{eq_var_stl}), we observe that the variance equation \eqref{eq_var_multiple} is always smaller because $W_t^{\top} (\cW \cW^{\top})^{-1} W_t \le 1$.
%On the other hand, the bias equation \eqref{eq_bias_multiple} is always larger because of the difference between the task models.
%We show the generalization error of hard parameter sharing estimators.
%Before stating the result, we define the following notations.

\medskip
\noindent\textbf{Setting.} Suppose we have $t$ datasets whose sample sizes are all equal to $n$ and whose features are all denoted by $X \in \real^{n \times p}$. The label vector of the $i$-th task follows a linear model $Y^{(i)} = X \beta^{(i)} + \varepsilon^{(i)}$.
We assume:
\begin{enumerate}
	\item $X = Z \Sigma^{1/2} \in \real^{n\times p}$ for a positive semidefinite matrix $\Sigma\in \R^{p\times p}$, and every entry of $Z \in \real^{n \times p}$ is drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$.
	\item every entry of $\varepsilon^{(i)} \in \real^{n \times t}$ is drawn indepdently from a one dimensional distribution with zero mean, variance $\sigma^2$, and bounded moment up to any order.\footnote{There exists a fixed function $C: \mathbb{N} \rightarrow \real^+$ such that for any $k \in \N$, the $k$-th moment is bounded by $C(k)$.}
\end{enumerate}
For an estimator $\hat{\beta}_i$ of task $i$, we are interested in its (out-of-sample) prediction loss
\[ L(\hat{\beta}_i) = \bignorm{\Sigma^{1/2} (\hat{\beta_i} - \beta^{(i)})}^2. \]
Recall that $r$ is the width of $B$.
We focus on cases where $r < t$, because otherwise the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \citet{WZR20}).

Our first main result shows that hard parameter sharing essentially approximates all tasks through a rank-$r$ subspace.
To formalize this geometric intuition, we introduce the matrix $B^\star \define [{\beta}^{(1)},{\beta}^{(2)},\dots,{\beta}^{(t)}] \in \real^{p\times t}$ which contains all the linear model parameters.
Let $A^{\star} {A^{\star}}^{\top}$ denote the best rank-$r$ subspace approximation of ${B^{\star}}^\top\Sigma B^{\star}$ (which is task labels' ``covariance''):\footnote{To ensure that $A^{\star}$ is unique, we assume that $\lambda_{r+1}({B^\star}^\top \Sigma B^\star)$ is strictly smaller than $\lambda_{r}({B^\star}^\top \Sigma B^\star)$.}
\begin{align}\label{eq_A_star}
	A^{\star} \define \argmin_{U\in\real^{t\times r} : U^{\top} U = \id_{r\times r}} \inner{U U^{\top}} {{B^{\star}}^{\top} \Sigma B^{\star}}.
\end{align}
Let $a_i^{\star} \in\real^r$ denote the $i$-th column of $A^{\star}{A^{\star}}^{\top}$.
We show that the prediction loss of HPS decomposes into a bias term $L(B^{\star} a_i^{\star})$ that measures the prediction loss of $B^{\star} a_i^{\star}$, plus a variance term that scales with $\norm{a_i^{\star}}^2$.
% $B^{\star} A^{\star} {A^{\star}}^{\top}$.
Let $(\hat{A}, \hat{B})$ be the global minimizer of $f(A, B)$.
Recall that the HPS estimator is defined as $\hat{\beta}_i^{\MTL} = \hat{B} \hat{A}_i$.
Our result is stated as follows.

@@

%\FY{why not use $C_1: = \frac{ \normFro{\Sigma^{1/2} B^{\star}}^2}{ \lambda_{\min}^2(\Sigma^{1/2} B^{\star}) + \sigma^2}$? $C_2$ does not seem to be correct because the dimension does not match. I will check the proof to see whether $C_2 =  C_1\cdot \norm{\Sigma^{1/2} B^{\star}}^2$ or $C_1\cdot \sigma\norm{\Sigma^{1/2} B^{\star}}$?}
\noindent\textbf{Comparison to single-task learning (STL).}
Theorem \ref{thm_many_tasks} provides a sharp generalization error bound that is asymptotically tight when $n$ goes to infinity.
%The limiting loss of hard parameter sharing consists of two parts, a bias term $L(B^{\star} a_i^{\star})$ that measures the error of $B^{\star} a_i^{\star}$, and a variance term that scales with noise variance $\sigma^2$.
%	Our result implies that the variance of hard parameter sharing is always smaller than single-task learning.
%	This is because	STL's variance is equal to $\frac{\sigma^2 \cdot p} {n - p}$ by Fact \ref{lem_minv}, and $\norm{a_i^{\star}}^2 \le 1$ since the spectral norm of $U_r$, which is a projection matrix, is at most one.
One direct implication of our result is that compared to STL, the variance always decreases, since STL's variance is equal to $\sigma^2 \tr[\Sigma (X^{\top} X)^{-1}]$.
On the other hand, the bias always increases.

\paragraph{How does hard parameter sharing scale with sample size $n$?}
Obviously, the concentration error decreases with $n$.
First, we consider the variance of $\hat{\beta}_i^{\MTL}$, which is $\sigma^2\norm{a_i^{\star}} \bigtr{\Sigma (X^{\top} X)^{-1}}$?
It turns out that this quantity converges to a fixed limit in the high-dimensional setting, which is formally stated in the following assumption.

@@

\begin{fact}[cf. Theorem 2.4 in \citet{isotropic}]\label{fact_tr}
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
		\[ \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} \pm \OO(n^{-c_{\varphi}}). \]
\end{fact}
\noindent\textit{Remark.} The above result has a long history in random matrix theory.
For a multivariate Gaussian random matrix, this result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, this result can be obtained using the well-known Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).
Applying Fact \ref{fact_tr} to Theorem \ref{thm_many_tasks}, we obtain that hard parameter sharing's variance is
		\[ \sigma^2 \norm{a^{\star}_i}^2 \bigtr{\Sigma (X^{\top} X)^{-1}} = \sigma^2 \norm{a^{\star}_i}^2 \frac{p}{n- p} \pm \OO(p^{- {c_{\varphi}} }). \]

@@

Next, we consider the bias of $\hat{\beta}_i^{\MTL}$, that is $L(B^{\star} a^{\star}_i)$.
We illustrate the bias through a random-effect model, which has been studied for a single-task case \cite{dobriban2020wonder}.
Suppose every $\beta^{(i)}$ consists of two random components, one that is shared among all tasks and one that is task-specific.
Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
Let $\beta_0$ denote the shared component whose entries are sampled i.i.d. from an isotropic Gaussian distribution of mean zero and variance $p^{-1}\kappa^2$.
Let $\beta^{(i)}$ be equal to $\beta_0$ plus a task-specific component that is a random Gaussian vector with i.i.d. entries of mean zero and variance $p^{-1} d^2$.
Thus, for any two different $\beta^{(i)}$ and $\beta^{(j)}$, their distance is roughly $2d^2$.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
Concretely, we can think of $\kappa = 1$ and $d^2/\sigma^2 = \OO(1)$.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%We assume that all the random variables have finite moments up to any order as in equation \eqref{assmAhigh2}.

>>>>

%We begin by considering the case where all tasks have the same sample size and feature covariates, that is, $n_i = n$ and $X^{(i)} = X\in\real^{n\times p}$ for all $i = 1, \dots, t$.
%We provide a sharp generalization error bound of hard parameter sharing estimators.
This section describes our result that the prediction loss of hard parameter sharing admits a clean bias-variance decomposition when all tasks have the same features.
%\FY{Is it should be "same covariate" or "same covariates"?}
%This setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
%We consider an arbitrary local minimum $B, W_1, \dots, W_2$ of the optimization objective.
%We extend the bias-variance decomposition from the two-task case to the multiple-task case.
%We observe that the expected prediction loss of $\hat{\beta}_t^{\MTL}$ conditional on $X$ consists of a bias and a variance equation as follows
%\begin{align}
%	\exarg{\varepsilon_1, \dots, \varepsilon_t}{L(\hat{\beta}_t^{\MTL}) \mid X}
%	=& \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} \cW^{\top} (\cW \cW^{\top})^{-1} W_t - \beta_t}}^2 \label{eq_bias_multiple} \\
%	&+ \sigma^2 \cdot (W_t^{\top} (\cW \cW^{\top})^{-1} W_t) \cdot \bigtr{\Sigma (X^{\top} X)^{-1}} \label{eq_var_multiple}
%\end{align}
%One can see that equation \eqref{eq_bias_multiple} is the bias of the multi-task learning estimator and equation \eqref{eq_var_multiple} is its variance.
%Compared to the prediction loss of single-task learning (cf. equation \eqref{eq_var_stl}), we observe that the variance equation \eqref{eq_var_multiple} is always smaller because $W_t^{\top} (\cW \cW^{\top})^{-1} W_t \le 1$.
%On the other hand, the bias equation \eqref{eq_bias_multiple} is always larger because of the difference between the task models.
%We show the generalization error of hard parameter sharing estimators.
%Before stating the result, we define the following notations.

\medskip
\noindent\textbf{Setting.} Suppose we have $t$ datasets whose sample sizes are all equal to $n$ and whose features are all denoted by $X \in \real^{n \times p}$. The label vector of the $i$-th task follows a linear model $Y^{(i)} = X \beta^{(i)} + \varepsilon^{(i)}$.
We assume:
\begin{enumerate}
	\item $X = Z \Sigma^{1/2} \in \real^{n\times p}$ for a positive semidefinite matrix $\Sigma\in \R^{p\times p}$, and every entry of $Z \in \real^{n \times p}$ is drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$.
	\item every entry of $\varepsilon^{(i)} \in \real^{n \times t}$ is drawn indepdently from a one dimensional distribution with zero mean, variance $\sigma^2$, and bounded moment up to any order.\footnote{There exists a fixed function $C: \mathbb{N} \rightarrow \real^+$ such that for any $k \in \N$, the $k$-th moment is bounded by $C(k)$.}
\end{enumerate}
For an estimator $\hat{\beta}_i$ of task $i$, we are interested in its (out-of-sample) prediction loss
\[ L(\hat{\beta}_i) = \bignorm{\Sigma^{1/2} (\hat{\beta_i} - \beta^{(i)})}^2. \]
Recall that $r$ is the width of $B$.
We focus on cases where $r < t$, because otherwise, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \citet{WZR20}).

Our first main result shows that hard parameter sharing essentially approximates all tasks through a rank-$r$ subspace.
To formalize this geometric intuition, we introduce the matrix $B^\star \define [{\beta}^{(1)},{\beta}^{(2)},\dots,{\beta}^{(t)}] \in \real^{p\times t}$, which contains all the linear model parameters.
Let $A^{\star} {A^{\star}}^{\top}$ denote the best rank-$r$ subspace approximation of ${B^{\star}}^\top\Sigma B^{\star}$ (which is task labels' ``covariance''):\footnote{To ensure that $A^{\star}$ is unique, we assume that $\lambda_{r+1}({B^\star}^\top \Sigma B^\star)$ is strictly smaller than $\lambda_{r}({B^\star}^\top \Sigma B^\star)$.}
\begin{align}\label{eq_A_star}
	A^{\star} \define \argmin_{U\in\real^{t\times r} : U^{\top} U = \id_{r\times r}} \inner{U U^{\top}} {{B^{\star}}^{\top} \Sigma B^{\star}}.
\end{align}
Let $a_i^{\star} \in\real^r$ denote the $i$-th column of $A^{\star}{A^{\star}}^{\top}$.
We show that the prediction loss of HPS decomposes into a bias term $L(B^{\star} a_i^{\star})$ that measures the prediction loss of $B^{\star} a_i^{\star}$, plus a variance term that scales with $\norm{a_i^{\star}}^2$.
% $B^{\star} A^{\star} {A^{\star}}^{\top}$.
Let $(\hat{A}, \hat{B})$ be the global minimizer of $f(A, B)$.
Recall that the HPS estimator is defined as $\hat{\beta}_i^{\MTL} = \hat{B} \hat{A}_i$.
Our result is stated as follows.

@@

%\FY{why not use $C_1: = \frac{ \normFro{\Sigma^{1/2} B^{\star}}^2}{ \lambda_{\min}^2(\Sigma^{1/2} B^{\star}) + \sigma^2}$? $C_2$ does not seem to be correct because the dimension does not match. I will check the proof to see whether $C_2 =  C_1\cdot \norm{\Sigma^{1/2} B^{\star}}^2$ or $C_1\cdot \sigma\norm{\Sigma^{1/2} B^{\star}}$?}
\noindent\textbf{Comparison to single-task learning (STL).}
Theorem \ref{thm_many_tasks} provides a sharp generalization error bound that is asymptotically tight when $n$ goes to infinity.
%The limiting loss of hard parameter sharing consists of two parts, a bias term $L(B^{\star} a_i^{\star})$ that measures the error of $B^{\star} a_i^{\star}$, and a variance term that scales with noise variance $\sigma^2$.
%	Our result implies that the variance of hard parameter sharing is always smaller than single-task learning.
%	This is because	STL's variance is equal to $\frac{\sigma^2 \cdot p} {n - p}$ by Fact \ref{lem_minv}, and $\norm{a_i^{\star}}^2 \le 1$ since the spectral norm of $U_r$, which is a projection matrix, is at most one.
One direct implication of our result is the variance always decreases when compared to STL, since STL's variance is equal to $\sigma^2 \tr[\Sigma (X^{\top} X)^{-1}]$.
On the other hand,  bias always increases.

\paragraph{How does hard parameter sharing scale with sample size $n$?}
Obviously, the concentration error decreases with $n$.
Therefore, we consider the bias and variance, respectively.
First, we consider the variance of $\hat{\beta}_i^{\MTL}$, which is $\sigma^2\norm{a_i^{\star}} \bigtr{\Sigma (X^{\top} X)^{-1}}$?
It turns out that this quantity converges to a fixed limit in the high-dimensional setting, which we define formally below.

@@

\begin{fact}[cf. Theorem 2.4 in \citet{isotropic}]\label{fact_tr}
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
		\[ \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} \pm \OO(n^{-c_{\varphi}}). \]
\end{fact}
\noindent\textit{Remark.} The above result has a long history in random matrix theory.
For a multivariate Gaussian random matrix, this result follows from the classical work for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For non-Gaussian random matrices, this result can be derived via the well-known Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).
Applying Fact \ref{fact_tr} to Theorem \ref{thm_many_tasks}, we obtain that hard parameter sharing's variance is
		\[ \sigma^2 \norm{a^{\star}_i}^2 \bigtr{\Sigma (X^{\top} X)^{-1}} = \sigma^2 \norm{a^{\star}_i}^2 \frac{p}{n- p} \pm \OO(p^{- {c_{\varphi}} }). \]

@@

Next, we consider the bias of $\hat{\beta}_i^{\MTL}$, which is $L(B^{\star} a^{\star}_i)$.
We illustrate the bias through a random-effect model, which has been studied for a single-task case \cite{dobriban2020wonder}.
Suppose every $\beta^{(i)}$ consists of two random components, one that is shared among all tasks and one that is task-specific.
Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
Let $\beta_0$ denote the shared component whose entries are sampled i.i.d. from an isotropic Gaussian distribution of mean zero and variance $p^{-1}\kappa^2$.
Let $\beta^{(i)}$ be equal to $\beta_0$ plus a task-specific component that is a random Gaussian vector with i.i.d. entries of mean zero and variance $p^{-1} d^2$.
Thus, for any two different $\beta^{(i)}$ and $\beta^{(j)}$, their distance is roughly $2d^2$.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
Concretely, we can think of $\kappa = 1$ and $d^2/\sigma^2 = \OO(1)$.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%We assume that all the random variables have finite moments up to any order as in equation \eqref{assmAhigh2}.
